Reintroduction of resistant frogs facilitates landscape-scale recovery in the presence of a lethal fungal disease

### 2024-09-27 doc/manuscript/translocation.tex
Use default cite method instead of natbib

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{blkarray}
\usepackage{lineno}\linenumbers
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{\textbf{Fig.}}
\else
  \newcommand\figurename{\textbf{Fig.}}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\usepackage{bookmark}

@@

The collection of data from live \emph{R. sierrae} described here
complies with all ethical regulations and was approved under the
University of California-Santa Barbara Institutional Animal Care and Use
Committe (IACUC) protocol number 478. For the 24 translocations we
conducted, we identified donor populations from which adult frogs
(\(\geq\) 40 mm snout-vent length) could be collected based on several
years of VES and skin swab collections \citep{knapp2016}, and results
from population genetic analyses \citep{poorten2017}. The populations
that we selected contained hundreds of \emph{R. sierrae} adults and
thousands of tadpoles. These relatively high abundances were the result
of recent increases following previous Bd-caused declines
\citep{knapp2016}. As is typical for recovering \emph{R. sierrae}
populations, Bd prevalence in the donor populations was high
(0.69--0.96) and Bd load (median log\textsubscript{10}(load) =
3.06--3.78 ITS copies) was two or more orders of magnitude below the
level at which frog mortality is expected
\citep{vredenburg2010, joseph2018} (log\textsubscript{10}(load)
\(\approx\) 5.78 ITS copies). Recipient sites to which frogs were
translocated were chosen based on previous \emph{R. sierrae} presence
(determined from VES and/or museum records) or characteristics that
suggested high quality habitat for this species \citep{knapp2005}. At
the beginning of this study, we had a relatively limited understanding
of the factors that affect habitat quality. In subsequent years, we
improved our site selection process by incorporating new information
about important habitat features, in particular, overwintering habitats
such as submerged boulders and overhanging banks. \emph{R. sierrae} were
absent from all recipient sites prior to the first translocation. This
determination was based on VES conducted at recipient sites over several
years (see additional details below regarding frog detectability during
VES).

@@

During each site visit, we conducted VES either immediately before CMR
surveys or during the first day of CMR surveys. VES was conducted by
walking the entire water body perimeter, first 100 m of each inlet and
outlet stream, and any fringing ponds and wetlands, and counting all
\emph{R. sierrae} tadpoles and juveniles. These \emph{R. sierrae} life
stages have high detectability, and counts are highly repeatable and
provide estimates of relative abundance \citep{knapp2000}.

@@

For each translocation site, we estimated survival of translocated
frogs, recruitment of new frogs into the adult population, and adult
population size using a site-specific Bayesian open-population
Jolly-Seber CMR model with known additions to the population (i.e.,
translocated cohorts), as implemented by the mrmr package
\citep{joseph2019} and using R Statistical Software
\citep{rsoftware2022} (v4.4.4) (see \textbf{Supplementary Methods - Frog
population recovery - CMR model structure} for details). Briefly, the
model tracks the states of \emph{M} individuals that comprise a
superpopulation made up of real and pseudo-individuals
\citep{joseph2018}. The possible states of individuals include ``not
recruited'', ``alive'', and ``dead''. The possible observations of
individuals include ``detected'' and ``not detected''. We assume that
individuals that are in the ``not recruited'' or ``dead'' states are
never detected (i.e., there are no mistakes in the individual PIT tag
records). We also assume that new recruits were the result of
within-site reproduction and not immigration from adjacent populations.
This assumption is justified by the fact that no \emph{R. sierrae}
populations were present within several kilometers of the translocation
sites. For all models, we used mrmr defaults for priors, number of
chains (4), and warmup and post-warmup iterations (2000 for each). We
evaluated convergence of the Markov chain Monte Carlo (MCMC) algorithm
using trace plots and Gelman-Rubin statistics (Rhat).

@@

To identify important predictors of frog survival following
translocation, we used multilevel Bayesian models
\citep{gelman2013, gabry2019}. Included predictor variables describe
characteristics of sites, translocated cohorts, and individuals (Bd
load, sex, frog size, site elevation, winter severity in the year of
translocation, winter severity in the year following translocation,
donor population, day of year on which a translocation was conducted,
and translocation order). We used 1-year post-translocation survival
estimates from CMR models as the response. Estimated survival was
rounded to integer values to produce a binary outcome, and modeled with
a Bernoulli distribution. Group-level (random) effects included
site\_id, translocation\_id, or translocation\_id nested within
site\_id. We performed all analyses with the rstanarm package
\citep{rstanarm2022} and R Statistical Software \citep{rsoftware2022}
(v4.4.4). For all models, we used default, weakly informative priors,
four chains, and 5000 iterations each for warmup and post-warmup. We
checked MCMC convergence using trace plots and Rhat, and evaluated model
fit using leave-one-out cross-validation \citep{vehtari2016}, as
implemented by the loo package \citep{vehtari2022}. (See
\textbf{Supplementary Methods - Frog population recovery - Among-site
survival modeling} for details.)

@@

We analyzed skin swabs using standard Bd DNA extraction and qPCR methods
\citep{boyle2004} (see \textbf{Supplementary Information - Frog
population recovery - Laboratory methods} for details). To assess the
magnitude of changes in Bd load on frogs following translocation, we
compared Bd loads measured before versus after translocation.
Before-translocation loads were quantified using skin swabs collected
from all to-be-translocated frogs at the donor site on the day before or
the day of the translocation. After-translocation Bd loads were based on
all swabs collected from translocated frogs at the recipient site in the
year of and the year following translocation. Individual frogs and their
associated Bd loads were included in the dataset only if frogs were
captured at the recipient site at least once during the 1-year period
following translocation.

@@

In this model we ignore density-dependent recruitment because we were
interested in the growth of the population from an initial
reintroduction and whether this growth was sufficient to prevent
extinction over 50 years following the introduction. We also did not
directly consider the dynamics of Bd in this model. We made this
decision because (i) translocated populations are infected with Bd at
high prevalence \citep{joseph2018}, and (ii) host density does not seem
to play a significant role in multi-year Bd infection dynamics in this
system \citep{wilber2022}. Thus, ignoring Bd infection dynamics and
instead assuming all host vital rates are in the presence of high Bd
prevalence significantly simplifies the model without much loss of
realism. Additional details are provided in \textbf{Supplementary
Methods - Population viability modeling - Incorporating yearly
variability in survival rates} and \textbf{Estimating model parameters}.

@@

The data used in this study are available at
\url{https://doi.org/10.5281/zenodo.13851590} \citep{knapp2024a}. This
includes the estimated survival of each reintroduced frog cohort, data
used to identify predictors of 1-year frog survival, data used to
estimate growth rates and extinction probabilities of reintroduced
populations, Bd loads on frogs before and after translocation, and
survival probabilities of translocated versus naturally recruited
adults. For the relevant figures, source data are provided in a Source
Data file. The CMR datasets from each reintroduction site that we used
to calculate post-translocation frog survival are available at
\url{https://doi.org/10.5281/zenodo.13851652} \citep{knapp2024b}.

@@

R code to replicate our analyses is available at
\url{https://doi.org/10.5281/zenodo.13851590} \citep{knapp2024a}. R code
to analyze the CMR datasets is available at
\url{https://doi.org/10.5281/zenodo.13851652} \citep{knapp2024b}.

@@

\renewcommand{\bibsection}{}
\bibliography{translocation.bib}

>>>>

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{blkarray}
\usepackage{lineno}\linenumbers
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{\textbf{Fig.}}
\else
  \newcommand\figurename{\textbf{Fig.}}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

The collection of data from live \emph{R. sierrae} described here
complies with all ethical regulations and was approved under the
University of California-Santa Barbara Institutional Animal Care and Use
Committe (IACUC) protocol number 478. For the 24 translocations we
conducted, we identified donor populations from which adult frogs
(\(\geq\) 40 mm snout-vent length) could be collected based on several
years of VES and skin swab collections\textsuperscript{16}, and results
from population genetic analyses\textsuperscript{54}. The populations
that we selected contained hundreds of \emph{R. sierrae} adults and
thousands of tadpoles. These relatively high abundances were the result
of recent increases following previous Bd-caused
declines\textsuperscript{16}. As is typical for recovering \emph{R.
sierrae} populations, Bd prevalence in the donor populations was high
(0.69--0.96) and Bd load (median log\textsubscript{10}(load) =
3.06--3.78 ITS copies) was two or more orders of magnitude below the
level at which frog mortality is expected\textsuperscript{31,39}
(log\textsubscript{10}(load) \(\approx\) 5.78 ITS copies). Recipient
sites to which frogs were translocated were chosen based on previous
\emph{R. sierrae} presence (determined from VES and/or museum records)
or characteristics that suggested high quality habitat for this
species\textsuperscript{55}. At the beginning of this study, we had a
relatively limited understanding of the factors that affect habitat
quality. In subsequent years, we improved our site selection process by
incorporating new information about important habitat features, in
particular, overwintering habitats such as submerged boulders and
overhanging banks. \emph{R. sierrae} were absent from all recipient
sites prior to the first translocation. This determination was based on
VES conducted at recipient sites over several years (see additional
details below regarding frog detectability during VES).

@@

During each site visit, we conducted VES either immediately before CMR
surveys or during the first day of CMR surveys. VES was conducted by
walking the entire water body perimeter, first 100 m of each inlet and
outlet stream, and any fringing ponds and wetlands, and counting all
\emph{R. sierrae} tadpoles and juveniles. These \emph{R. sierrae} life
stages have high detectability, and counts are highly repeatable and
provide estimates of relative abundance\textsuperscript{29}.

@@

For each translocation site, we estimated survival of translocated
frogs, recruitment of new frogs into the adult population, and adult
population size using a site-specific Bayesian open-population
Jolly-Seber CMR model with known additions to the population (i.e.,
translocated cohorts), as implemented by the mrmr
package\textsuperscript{43} and using R Statistical
Software\textsuperscript{56} (v4.4.4) (see \textbf{Supplementary Methods
- Frog population recovery - CMR model structure} for details). Briefly,
the model tracks the states of \emph{M} individuals that comprise a
superpopulation made up of real and
pseudo-individuals\textsuperscript{39}. The possible states of
individuals include ``not recruited'', ``alive'', and ``dead''. The
possible observations of individuals include ``detected'' and ``not
detected''. We assume that individuals that are in the ``not recruited''
or ``dead'' states are never detected (i.e., there are no mistakes in
the individual PIT tag records). We also assume that new recruits were
the result of within-site reproduction and not immigration from adjacent
populations. This assumption is justified by the fact that no \emph{R.
sierrae} populations were present within several kilometers of the
translocation sites. For all models, we used mrmr defaults for priors,
number of chains (4), and warmup and post-warmup iterations (2000 for
each). We evaluated convergence of the Markov chain Monte Carlo (MCMC)
algorithm using trace plots and Gelman-Rubin statistics (Rhat).

@@

To identify important predictors of frog survival following
translocation, we used multilevel Bayesian
models\textsuperscript{57,58}. Included predictor variables describe
characteristics of sites, translocated cohorts, and individuals (Bd
load, sex, frog size, site elevation, winter severity in the year of
translocation, winter severity in the year following translocation,
donor population, day of year on which a translocation was conducted,
and translocation order). We used 1-year post-translocation survival
estimates from CMR models as the response. Estimated survival was
rounded to integer values to produce a binary outcome, and modeled with
a Bernoulli distribution. Group-level (random) effects included
site\_id, translocation\_id, or translocation\_id nested within
site\_id. We performed all analyses with the rstanarm
package\textsuperscript{59} and R Statistical
Software\textsuperscript{56} (v4.4.4). For all models, we used default,
weakly informative priors, four chains, and 5000 iterations each for
warmup and post-warmup. We checked MCMC convergence using trace plots
and Rhat, and evaluated model fit using leave-one-out
cross-validation\textsuperscript{60}, as implemented by the loo
package\textsuperscript{61}. (See \textbf{Supplementary Methods - Frog
population recovery - Among-site survival modeling} for details.)

@@

We analyzed skin swabs using standard Bd DNA extraction and qPCR
methods\textsuperscript{62} (see \textbf{Supplementary Information -
Frog population recovery - Laboratory methods} for details). To assess
the magnitude of changes in Bd load on frogs following translocation, we
compared Bd loads measured before versus after translocation.
Before-translocation loads were quantified using skin swabs collected
from all to-be-translocated frogs at the donor site on the day before or
the day of the translocation. After-translocation Bd loads were based on
all swabs collected from translocated frogs at the recipient site in the
year of and the year following translocation. Individual frogs and their
associated Bd loads were included in the dataset only if frogs were
captured at the recipient site at least once during the 1-year period
following translocation.

@@

In this model we ignore density-dependent recruitment because we were
interested in the growth of the population from an initial
reintroduction and whether this growth was sufficient to prevent
extinction over 50 years following the introduction. We also did not
directly consider the dynamics of Bd in this model. We made this
decision because (i) translocated populations are infected with Bd at
high prevalence\textsuperscript{39}, and (ii) host density does not seem
to play a significant role in multi-year Bd infection dynamics in this
system\textsuperscript{63}. Thus, ignoring Bd infection dynamics and
instead assuming all host vital rates are in the presence of high Bd
prevalence significantly simplifies the model without much loss of
realism. Additional details are provided in \textbf{Supplementary
Methods - Population viability modeling - Incorporating yearly
variability in survival rates} and \textbf{Estimating model parameters}.

@@

The data used in this study are available at
\url{https://doi.org/10.5281/zenodo.13851590}\textsuperscript{64}. This
includes the estimated survival of each reintroduced frog cohort, data
used to identify predictors of 1-year frog survival, data used to
estimate growth rates and extinction probabilities of reintroduced
populations, Bd loads on frogs before and after translocation, and
survival probabilities of translocated versus naturally recruited
adults. For the relevant figures, source data are provided in a Source
Data file. The CMR datasets from each reintroduction site that we used
to calculate post-translocation frog survival are available at
\url{https://doi.org/10.5281/zenodo.13851652}\textsuperscript{65}.

@@

R code to replicate our analyses is available at
\url{https://doi.org/10.5281/zenodo.13851590}\textsuperscript{64}. R
code to analyze the CMR datasets is available at
\url{https://doi.org/10.5281/zenodo.13851652}\textsuperscript{65}.

@@

\phantomsection\label{refs}
\begin{CSLReferences}{0}{0}
\bibitem[\citeproctext]{ref-ceballos2015}
\CSLLeftMargin{1. }%
\CSLRightInline{Ceballos, G. \emph{et al.} Accelerated modern
human-induced species losses: Entering the sixth mass extinction.
\emph{Science Advances} \textbf{1}, (2015).}

\bibitem[\citeproctext]{ref-naeem2009}
\CSLLeftMargin{2. }%
\CSLRightInline{Naeem, S., Bunker, D. E., Hector, A., Loreau, M. \&
Perrings, C. \emph{Biodiversity, Ecosystem Functioning, and Human
Wellbeing: An Ecological and Economic Perspective}. (Oxford University
Press, Oxford, 2009).}

\bibitem[\citeproctext]{ref-jones2008}
\CSLLeftMargin{3. }%
\CSLRightInline{Jones, K. E. \emph{et al.} Global trends in emerging
infectious diseases. \emph{Nature} \textbf{451}, 990--993 (2008).}

\bibitem[\citeproctext]{ref-fisher2012}
\CSLLeftMargin{4. }%
\CSLRightInline{Fisher, M. C. \emph{et al.} Emerging fungal threats to
animal, plant and ecosystem health. \emph{Nature} \textbf{484}, 186--94
(2012).}

\bibitem[\citeproctext]{ref-daszak2000}
\CSLLeftMargin{5. }%
\CSLRightInline{Daszak, P., Cunningham, A. A. \& Hyatt, A. D. Emerging
infectious diseases of wildlife -- threats to biodiversity and human
health. \emph{Science} \textbf{287}, 443--449 (2000).}

\bibitem[\citeproctext]{ref-hewson2014}
\CSLLeftMargin{6. }%
\CSLRightInline{Hewson, I. \emph{et al.} Densovirus associated with
sea-star wasting disease and mass mortality. \emph{Proceedings of the
National Academy of Sciences USA} \textbf{111}, 17278--17283 (2014).}

\bibitem[\citeproctext]{ref-samuel2015}
\CSLLeftMargin{7. }%
\CSLRightInline{Samuel, M. D., Woodworth, B. L., Atkinson, C. T., Hart,
P. J. \& LaPointe, D. A. Avian malaria in {H}awaiian forest birds:
Infection and population impacts across species and elevations.
\emph{Ecosphere} \textbf{6}, art104 (2015).}

\bibitem[\citeproctext]{ref-scheele2019}
\CSLLeftMargin{8. }%
\CSLRightInline{Scheele, B. C. \emph{et al.} Amphibian fungal panzootic
causes catastrophic and ongoing loss of biodiversity. \emph{Science}
\textbf{363}, 1459--1463 (2019).}

\bibitem[\citeproctext]{ref-cunningham2021}
\CSLLeftMargin{9. }%
\CSLRightInline{Cunningham, C. X. \emph{et al.} Quantifying 25 years of
disease-caused declines in {T}asmanian devil populations: {h}ost density
drives spatial pathogen spread. \emph{Ecology Letters} \textbf{24},
958--969 (2021).}

\bibitem[\citeproctext]{ref-luedtke2023}
\CSLLeftMargin{10. }%
\CSLRightInline{Luedtke, J. A. \emph{et al.} Ongoing declines for the
world{'}s amphibians in the face of emerging threats. \emph{Nature}
\textbf{622}, 308--314 (2023).}

\bibitem[\citeproctext]{ref-raberg2009}
\CSLLeftMargin{11. }%
\CSLRightInline{Råberg, L., Graham, A. L. \& Read, A. F. Decomposing
health: Tolerance and resistance to parasites in animals.
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences} \textbf{364}, 37--49 (2008).}

\bibitem[\citeproctext]{ref-brannelly2021}
\CSLLeftMargin{12. }%
\CSLRightInline{Brannelly, L. A. \emph{et al.} Mechanisms underlying
host persistence following amphibian disease emergence determine
appropriate management strategies. \emph{Ecology Letters} \textbf{24},
130--148 (2021).}

\bibitem[\citeproctext]{ref-garner2016}
\CSLLeftMargin{13. }%
\CSLRightInline{Garner, T. W. J. \emph{et al.} Mitigating amphibian
chytridiomycoses in nature. \emph{Philosophical Transactions of the
Royal Society B: Biological Sciences} \textbf{371}, 20160207 (2016).}

\bibitem[\citeproctext]{ref-scheele2017}
\CSLLeftMargin{14. }%
\CSLRightInline{Scheele, B. C. \emph{et al.} After the epidemic: Ongoing
declines, stabilizations and recoveries in amphibians afflicted by
chytridiomycosis. \emph{Biological Conservation} \textbf{206}, 37--46
(2017).}

\bibitem[\citeproctext]{ref-voyles2018}
\CSLLeftMargin{15. }%
\CSLRightInline{Voyles, J. \emph{et al.} Shifts in disease dynamics in a
tropical amphibian assemblage are not due to pathogen attenuation.
\emph{Science} \textbf{359}, 1517--1519 (2018).}

\bibitem[\citeproctext]{ref-knapp2016}
\CSLLeftMargin{16. }%
\CSLRightInline{Knapp, R. A. \emph{et al.} Large-scale recovery of an
endangered amphibian despite ongoing exposure to multiple stressors.
\emph{Proceedings of the National Academy of Sciences USA} \textbf{113},
11889--11894 (2016).}

\bibitem[\citeproctext]{ref-rosenblum2012}
\CSLLeftMargin{17. }%
\CSLRightInline{Rosenblum, E. B., Poorten, T. J., Settles, M. \&
Murdoch, G. K. Only skin deep: Shared genetic response to the deadly
chytrid fungus in susceptible frog species. \emph{Molecular Ecology}
\textbf{21}, 3110--3120 (2012).}

\bibitem[\citeproctext]{ref-fites2013}
\CSLLeftMargin{18. }%
\CSLRightInline{Fites, J. S. \emph{et al.} The invasive chytrid fungus
of amphibians paralyzes lymphocyte responses. \emph{Science}
\textbf{342}, 366369 (2013).}

\bibitem[\citeproctext]{ref-grogan2018a}
\CSLLeftMargin{19. }%
\CSLRightInline{Grogan, L. F. \emph{et al.} Review of the amphibian
immune response to chytridiomycosis, and future directions.
\emph{Frontiers in Immunology} \textbf{9}, 2536 (2018).}

\bibitem[\citeproctext]{ref-savage2016}
\CSLLeftMargin{20. }%
\CSLRightInline{Savage, A. E. \& Zamudio, K. R. Adaptive tolerance to a
pathogenic fungus drives major histocompatibility complex evolution in
natural amphibian populations. \emph{Proceedings of the Royal Society B:
Biological Sciences} \textbf{283}, 20153115 (2016).}

\bibitem[\citeproctext]{ref-grogan2018b}
\CSLLeftMargin{21. }%
\CSLRightInline{Grogan, L. F. \emph{et al.} Evolution of resistance to
chytridiomycosis is associated with a robust early immune response.
\emph{Molecular Ecology} \textbf{27}, 919--934 (2018).}

\bibitem[\citeproctext]{ref-hammond2021}
\CSLLeftMargin{22. }%
\CSLRightInline{Hammond, T. T. \emph{et al.} Overwinter behavior,
movement, and survival in a recently reintroduced, endangered amphibian,
\emph{{R}ana muscosa}. \emph{Journal for Nature Conservation}
\textbf{64}, 126086 (2021).}

\bibitem[\citeproctext]{ref-knapp2022}
\CSLLeftMargin{23. }%
\CSLRightInline{Knapp, R. A. \emph{et al.} Effectiveness of antifungal
treatments during chytridiomycosis epizootics in populations of an
endangered frog. \emph{PeerJ} \textbf{10}, e12712 (2022).}

\bibitem[\citeproctext]{ref-stockwell2008}
\CSLLeftMargin{24. }%
\CSLRightInline{Stockwell, M., Clulow, S., Clulow, J. \& Mahony, M. The
impact of the {Amphibian Chytrid Fungus} on a {Green and Golden Bell
Frog} {\emph{{L}itoria aurea}} reintroduction program at the {Hunter
Wetlands Centre Australia in the Hunter Region of NSW}. \emph{Australian
Zoologist} \textbf{34}, 379--386 (2008).}

\bibitem[\citeproctext]{ref-scheele2021}
\CSLLeftMargin{25. }%
\CSLRightInline{Scheele, B. C. \emph{et al.} Conservation translocations
for amphibian species threatened by chytrid fungus: A review, conceptual
framework, and recommendations. \emph{Conservation Science and Practice}
\textbf{3}, e524 (2021).}

\bibitem[\citeproctext]{ref-grinnell1924}
\CSLLeftMargin{26. }%
\CSLRightInline{Grinnell, J. \& Storer, T. I. \emph{{A}nimal {L}ife in
the {Y}osemite}. (University of California Press, Berkeley, 1924).}

\bibitem[\citeproctext]{ref-vredenburg2007}
\CSLLeftMargin{27. }%
\CSLRightInline{Vredenburg, V. T. \emph{et al.} Concordant molecular and
phenotypic data delineate new taxonomy and conservation priorities for
the endangered mountain yellow-legged frog. \emph{Journal of Zoology}
\textbf{271}, 361--374 (2007).}

\bibitem[\citeproctext]{ref-bradford1989}
\CSLLeftMargin{28. }%
\CSLRightInline{Bradford, D. F. Allotopic distribution of native frogs
and introduced fishes in high {S}ierra {N}evada lakes of {C}alifornia:
{i}mplication of the negative effect of fish introductions.
\emph{Copeia} \textbf{1989}, 775--778 (1989).}

\bibitem[\citeproctext]{ref-knapp2000}
\CSLLeftMargin{29. }%
\CSLRightInline{Knapp, R. A. \& Matthews, K. R. Non-native fish
introductions and the decline of the mountain yellow-legged frog from
within protected areas. \emph{Conservation Biology} \textbf{14},
428--438 (2000).}

\bibitem[\citeproctext]{ref-vredenburg2019}
\CSLLeftMargin{30. }%
\CSLRightInline{Vredenburg, V. T. \emph{et al.} Pathogen invasion
history elucidates contemporary host pathogen dynamics. \emph{PLOS ONE}
\textbf{14}, e0219981 (2019).}

\bibitem[\citeproctext]{ref-vredenburg2010}
\CSLLeftMargin{31. }%
\CSLRightInline{Vredenburg, V. T., Knapp, R. A., Tunstall, T. S. \&
Briggs, C. J. Dynamics of an emerging disease drive large-scale
amphibian population extinctions. \emph{Proceedings of the National
Academy of Sciences USA} \textbf{107}, 9689--9694 (2010).}

\bibitem[\citeproctext]{ref-rachowicz2006}
\CSLLeftMargin{32. }%
\CSLRightInline{Rachowicz, L. J. \emph{et al.} Emerging infectious
disease as a proximate cause of amphibian mass mortality. \emph{Ecology}
\textbf{87}, 1671--1683 (2006).}

\bibitem[\citeproctext]{ref-knapp1998}
\CSLLeftMargin{33. }%
\CSLRightInline{Knapp, R. A. \& Matthews, K. R. Eradication of nonnative
fish by gill netting from a small mountain lake in {C}alifornia.
\emph{Restoration Ecology} \textbf{6}, 207213 (1998).}

\bibitem[\citeproctext]{ref-knapp2007}
\CSLLeftMargin{34. }%
\CSLRightInline{Knapp, R. A., Boiano, D. M. \& Vredenburg, V. T. Removal
of nonnative fish results in population expansion of a declining
amphibian (mountain yellow-legged frog, \emph{{R}ana muscosa}).
\emph{Biological Conservation} \textbf{135}, 11--20 (2007).}

\bibitem[\citeproctext]{ref-vredenburg2004}
\CSLLeftMargin{35. }%
\CSLRightInline{Vredenburg, V. T. Reversing introduced species effects:
Experimental removal of introduced fish leads to rapid recovery of a
declining frog. \emph{Proceedings of the National Academy of Sciences
USA} \textbf{101}, 7646--7650 (2004).}

\bibitem[\citeproctext]{ref-walker2007}
\CSLLeftMargin{36. }%
\CSLRightInline{Walker, S. \emph{et al.} Environmental detection of
\emph{{B}atrachochytrium dendrobatidis} in a temperate climate.
\emph{Diseases of Aquatic Organisms} \textbf{77}, 105--112 (2007).}

\bibitem[\citeproctext]{ref-briggs2010}
\CSLLeftMargin{37. }%
\CSLRightInline{Briggs, C. J., Knapp, R. A. \& Vredenburg, V. T.
Enzootic and epizootic dynamics of the chytrid fungal pathogen of
amphibians. \emph{Proceedings of the National Academy of Sciences USA}
\textbf{107}, 9695--9700 (2010).}

\bibitem[\citeproctext]{ref-knapp2011}
\CSLLeftMargin{38. }%
\CSLRightInline{Knapp, R. A., Briggs, C. J., Smith, T. C. \& Maurer, J.
R. Nowhere to hide: Impact of a temperature-sensitive amphibian pathogen
along an elevation gradient in the temperate zone. \emph{Ecosphere}
\textbf{2}, art93 (2011).}

\bibitem[\citeproctext]{ref-joseph2018}
\CSLLeftMargin{39. }%
\CSLRightInline{Joseph, M. B. \& Knapp, R. A. Disease and climate
effects on individuals drive post{-}reintroduction population dynamics
of an endangered amphibian. \emph{Ecosphere} \textbf{9}, e02499 (2018).}

\bibitem[\citeproctext]{ref-carlson2014}
\CSLLeftMargin{40. }%
\CSLRightInline{Carlson, S. M., Cunningham, C. J. \& Westley, P. A.
Evolutionary rescue in a changing world. \emph{Trends in Ecology \&
Evolution} \textbf{29}, 521--530 (2014).}

\bibitem[\citeproctext]{ref-searle2020}
\CSLLeftMargin{41. }%
\CSLRightInline{Searle, C. L. \& Christie, M. R. Evolutionary rescue in
host-pathogen systems. \emph{Evolution} \textbf{75}, 2948--2958 (2021).}

\bibitem[\citeproctext]{ref-mendelson2019}
\CSLLeftMargin{42. }%
\CSLRightInline{Mendelson III, J. R., Whitfield, S. M. \& Sredl, M. J.
{A recovery engine strategy for amphibian conservation in the context of
disease}. \emph{Biological Conservation} \textbf{236}, 188--191 (2019).}

\bibitem[\citeproctext]{ref-joseph2019}
\CSLLeftMargin{43. }%
\CSLRightInline{Joseph, M. B. {mrmr: mark recapture miscellany in R}.
(2019).}

\bibitem[\citeproctext]{ref-dodd2005}
\CSLLeftMargin{44. }%
\CSLRightInline{Dodd Jr, C. Amphibian conservation and population
manipulation. in \emph{Status and conservation of {US} amphibians} (ed.
Lannoo, M.) 265--270 (University of California Press, Berkeley, 2005).}

\bibitem[\citeproctext]{ref-seddon2007}
\CSLLeftMargin{45. }%
\CSLRightInline{Seddon, P. J., Armstrong, D. P. \& Maloney, R. F.
Developing the science of reintroduction biology. \emph{Conservation
Biology} \textbf{21}, 303--312 (2007).}
